---
title: "Homework 2"
author: Ling Tuo
date: 09/29/2020
output: 
  github_document:
    toc: true
---

This is my solution to HW2. 

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./p1p2_datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
        "./p1p2_datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
        sheet = "2018 Precipitation",
        skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
        "./p1p2_datasets/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
        sheet = "2017 Precipitation",
        skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

```

Now combine annual precipitation.

```{r, eval = FALSE}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.

The total precipitation in 2018 is `r sum(pull(precip_2018,total))` and the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trashwheel_df,year=="2017"), sports_balls))`.


## Problem 2

Read and clean the NYC Transit dataset. 

```{r, message = FALSE}
nyctransit_df = 
  read_csv(
    "./p1p2_datasets/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    ada_compliance = ada
    ) %>% 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada_compliance
    ) %>% 
  mutate(
    entry = case_when(entry == "YES" ~ "TRUE", entry == "NO" ~ "FALSE"), 
    entry = as.logical(entry)
    )
```

- This dataset contains information of the subway transformation in NYC, related to entrance and exit for each station. The dataset contains information on line, station name and location (longitude, latitude), routes served, entry, vending, entrance type, and ADA compliance. 

- Firstly, I load the dataset into nyctransit_df, clean the variable names and rename the 'ada' and select variables of interest.

- Then I convert the entry variable from character (YES vs NO) to a logical variable(TRUE vs FALSE). It is not a completely tidy format, since there are several columns of "routes_served" that can be addressed into more clear format (in the following text).

- To sum up, there are a total of `r nrow(nyctransit_df)` rows, `r ncol(nyctransit_df)` columns in the final dataset.

Other related questions:

- There are `r nyctransit_df %>% distinct(station_name, line) %>% nrow` distinct stations, identified both by name and by line.

- There are `r nyctransit_df %>% filter(ada_compliance == "TRUE") %>% nrow` stations are ADA compliant.

-  The proportion of station entrances / exits without vending allow entrance is `r  nyctransit_df %>% filter(vending == "NO" & entry == "TRUE") %>% nrow / nyctransit_df %>% filter(vending == "NO") %>% nrow`

Reformat data so that route number and route name are distinct variables: 

```{r}
nyctransit_df_tidy = 
  nyctransit_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "routes_name",
    names_prefix = "route",
    values_to = "routes_number"
    ) %>% 
  drop_na(routes_number) 
```

- There are  `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A") %>% nrow`  distinct stations serve the A train.

- Of the stations that serve the A train, there are `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A" & ada_compliance == "TRUE") %>% nrow` ADA compliant.

## Problem 3

Read and clean pols-month dataset.

```{r, message = FALSE}
pols_month_df = 
  read_csv(
    "./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  arrange(year, month) %>% 
  mutate_at(vars(year:day), as.integer) %>% 
  mutate(
    month = recode(month, `1` = "January", `2` = "February", `3` = "March", `4` = "April", `5` = "May", `6` = "June", `7` = "July", `8` = "August", `9` = "September", `10` = "October", `11` = "November", `12` = "December")
    ) %>% 
  mutate(
    president = case_when(prez_gop == "1" ~ "gop", prez_dem == "1" ~ "dem")
    ) %>% 
  select(-prez_dem, -prez_gop)
```

Read and clean snp dataset.

```{r, message = FALSE}
snp_df = 
  read_csv(
    "./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  select(year, month, day, everything()) %>% 
  mutate_at(vars(year:day), as.integer) %>% 
  arrange(year, month) %>%
  mutate(
    month = recode(month, `1` = "January", `2` = "February", `3` = "March", `4` = "April", `5` = "May", `6` = "June", `7` = "July", `8` = "August", `9` = "September", `10` = "October", `11` = "November", `12` = "December")
    ) 
```

Read and clean unemployment dataset.

```{r, message = FALSE}
unemployment_df = 
  read_csv(
    "./fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage"
    ) %>% 
  arrange(year) %>%
  mutate(
    month = recode(month, `jan` = "January", `feb` = "February", `mar` = "March", `apr` = "April", `may` = "May", `jun` = "June", `jul` = "July", `aug` = "August", `sep` = "September", `oct` = "October", `nov` = "November", `dec` = "December")
    ) 
```

Merging snp into pols and merging unemployment into the result.

```{r}
pols_snp = left_join(pols_month_df, snp_df, by = c("year", "month"))

pols_snp_unemploy = left_join(pols_snp, unemployment_df, by = c("year", "month"))
```

The pols-month dataset contains information of the number of national politicians who are democratic or republican at any given time. The dataset contains year, month, day, the number of republican or democratic governors/senators/representatives. 

The snp dataset contains information of closing values of Standard & Poor’s stock market index (S&P), including variables of date and close.

The unemployment dataset contains information of percentage of unemployment in different month of a associated year, including variables of year, month, percentage (after adjusted).

As for the final dataset, there are `r nrow(pols_snp_unemploy)` rows and `r ncol(pols_snp_unemploy)` columns. It combines above three datasets including variables of year, month, day, president, percentage of unemployment, close, etc. The range of years is [`r range(pull(pols_snp_unemploy, year))`].
